Relationships in the extremes and their influence on competition and coexistence: Supporting information

### 2021-10-28 paper/SupMat.Rnw
finished revising notation in SupMat

\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))] - \E[\ln(1-\delta+\delta\exp(b_i^\shp-b_j^\shp))],\label{eq:decomp:contrib_EC_1}

@@

\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))]
, \label{eq:decomp:contrib_EC}
$(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$ and $\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent.
Continuing, $\olep_i^{[EC]}=\olep_i^{(EC)}-\olep_i^{[E \nat C]}$ is available by subtracting expression (\ref{eq:decomp:contrib_EC})

@@

\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))],\label{eq:decomp:contrib_ATA_2}

@@

@@

\olep_j^{[E \nat C]} &= \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] - \E[r_j(E_j^\shp , C_{j \bs i}^\shp)]\\
&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],
\label{eq:decomp:j:EChard}
\olep_j^{[EC]} &= \olep_j^{(EC)} - \olep_j^{[E \nat C]} \\
&= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)]

@@

$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other.

@@

We now examine $\olep_i^{[EC ]}$, which, by (\ref{eq:decomp:contrib_ATA_2}), equals 
\begin{equation}
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
\end{equation}
But $(\ddot{b}_i,\ddot{b}_j)$ is a bivariate normal random variable constructed to have the same marginals and 
covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
$(b_i^{(r)},b_j^{(r)})$, so $(\ddot{b}_i^{(l)},\ddot{b}_j^{(l)})$ is distributed in the same way as 
$(\ddot{b}_i^{(r)},\ddot{b}_j^{(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
of the choice between our left- and right-tail-associated noise cases. Now examining 
$\E[\ln(1-\delta+\delta\exp(b_i-b_j))]$, we define $\beta_1=-b_2+\mu_1+\mu_2$ and $\beta_2=-b_1+\mu_1+\mu_2$,
and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
$(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
$(b_1^{(l)},b_2^{(l)})$. But then we have 
\begin{align}
\E[\ln(1-\delta+\delta\exp(b_i^{(l)}-b_j^{(l)}))] &= \E[\ln(1-\delta+\delta\exp(-\beta_j^{(l)}+\mu_1+\mu_2-(-\beta_i^{(l)}+\mu_1+\mu_2)))] \\
&= \E[\ln(1-\delta+\delta\exp(\beta_i^{(l)}-\beta_j^{(l)}))] \\
&= \E[\ln(1-\delta+\delta\exp(b_i^{(r)}-b_j^{(r)}))], 
\end{align}
which completes the proof that $\olep_i^{[EC]}$ is independent of left- versus right-tail associated noise.
For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 
above, in this section, and note that all those terms depend only on marginals.

@@

Next, given values of the parameters $\delta$, $\mu_i - \mu_j$, and $\sigma$, proceed as follows to get estimates of the 
terms of the decomposition (\ref{eq:decomp_overall_IGR}).
\begin{enumerate}
\item Compute $\epsilon_i^0$ following (\ref{eq:decomp:baseline}), which is a simple formula of $\delta$ and 
$\mu_i - \mu_j$, not requiring use of the random samples described above. This gives an exact value of
$\epsilon_i^0$, not an estimate.
\item Estimate $\olep_i^E$, following (\ref{eq:decomp:contrib_of_var_E}), as 
$\widehat{\olep_i^E}=\mean_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0$, an estimate with 
standard error $\se \left( \widehat{\olep_i^E} \right) = \sd_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]/\sqrt{M}$,
where $\sd(\cdot)$ denotes sample standard deviation.
\item Estimate $\olep_i^C$, following (\ref{eq:decomp:contrib_of_var_C}), as 
$\widehat{\olep_i^C}=\mean_k [\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-\epsilon_i^0$, which has 
standard error $\se \left( \widehat{\olep_i^C} \right) = \sd_k[\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_i^\shp-\tilde{b}_j^\shp$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) is normally distributed
with mean $0$ and variance $2$, so we can estimate $\olep_i^{(E \shp C)}$, following (\ref{eq:decomp:contrib_of_EC_var_per_se}),
as $\widehat{\olep_i^{(E \shp C)}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0.$ The standard error of this expression is 
$\se \left( \widehat{\olep_i^{(E \shp C)}} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma \sqrt{2} u^{(k)}+\mu_i-\mu_j))-
\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$. 
\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_EC}), as
$\widehat{\olep_i^{[EC]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$. The standard error of this expression is 
$\se \left( \widehat{\olep_i^{[EC]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$
and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_ATA_2}), as 
$\widehat{\olep_i^{[E \nat C]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]-
\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))].$ The standard
error of this expression is $\se \left( \widehat{\olep_i^{[E \nat C]}} \right)=\sqrt{s_3^2+s_1^2}$, where
$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
\item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:base_IGR}), as 
$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))],$
which has standard error $s_3$ from above.
\end{enumerate}
It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
\widehat{\olep_i^{[EC]}}+\widehat{\olep_i^{[E \nat C]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
at the level of the estimators, as well

Next, proceed as follows to get estimates of the terms in the decomposition (\ref{eq:decomp_overall_rjbsi}). 
\begin{enumerate}
\item Use $\epsilon_j^0=0$, as in (\ref{eq:decomp:baseline_j}), and this is the exact value.
\item Estimate $\olep_j^E$, following (\ref{eq:decomp:contrib_of_var_E_j}), as
$\widehat{\olep_j^E}=\mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]-\epsilon_j^0$, an estimate
with standard error $\se \left( \widehat{\olep_j^E} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\olep_j^C$, following (\ref{eq:decomp:contrib_of_var_C_j}), as
$\widehat{\olep_j^C}=\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]-\epsilon_j^0$, which has 
standard error $\se \left( \widehat{\olep_j^C} \right) = \sd_k [\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_j$ and $\tilde{b}_j^\shp$ in (\ref{eq:decomp:j:EshpC}) are standard normally distributed and
independent, so $\tilde{b}_j^\shp - \tilde{b}_j$ is normally distributed with mean $0$ and variance $2$. So we can 
estimate $\olep_j^{(E \shp C)}$, following (\ref{eq:decomp:j:EshpC}), as 
$\widehat{\olep_j^{(E \shp C)}} = \mean_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\olep_j^E}-\widehat{\olep_j^C}-\epsilon_j^0$.
The standard error of this expression is $\se \left(  \widehat{\olep_j^{(E \shp C)}} \right) = 
\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))-\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\olep_j^{[EC]}$, following (\ref{eq:decomp:j:EChard}), as $\widehat{\olep_j^{[EC]}}=\mean_k [-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))],$ which has $\se \left( \widehat{\olep_j^{[EC]}} \right) = \sd_k[-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))]/\sqrt{M}$.
\item Use $\olep_j^{[E \nat C]}=0$, following (\ref{eq:decomp:j:EnatC}). This is exact.
\item Finally, use $\olr_{j \bs i} = 0$, again exact. 
\end{enumerate}
As for (\ref{eq:decomp_overall_IGR}), it is again easy to see that (\ref{eq:decomp_overall_rjbsi}) applies for the 
estimates as well as for the quantities being estimated.

@@

\item Estimate $\Delta_i^{[EC]}$ as $\widehat{\Delta_i^{[EC]}}=\widehat{\olep_i^{[EC]}}-q_{ij}\widehat{\olep_j^{[EC]}}=

@@

\item For $\Delta_i^{[E \nat C]}$, we recall that $\olep_j^{[E \nat C]}=0$, so $\Delta_i^{[E \nat C]}=\olep_i^{[E \nat C]}$. So we use
$\widehat{\olep_i^{[E \nat C]}}$ for $\widehat{\Delta_i^{[E \nat C]}}$.

@@

In fig \ref{MT-fig:fig4} we plot IGR , $\Delta_i^{[E \nat C]}$, and IGR $-\Delta_i^{[E \nat C]}$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.  
From section \ref{sect:math_simplifications}, 
\begin{align}
\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
&= 0,
\end{align}
and
\begin{align}
\olep_j^{[E \nat C]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
&= 0. %\label{eq:decomp:j:EnatC}
\end{align}
Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[E \nat C]}$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$ are zero so 
\begin{equation}
\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\olr_{i \bs i}
\end{equation}
and 
\begin{equation}
\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}=\olep_i^{[E \nat C]}
\end{equation}
Similarly for IGR $-\Delta_i^{[E \nat C]}$, 
\begin{equation}
(\olr_{i \bs i}-q_{ij}\olr_{j \bs i})-\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]} = \olr_{i \bs i}-\olep_i^{[E \nat C]}.
\end{equation}
Thus $q_{ij}$ does not matter for anything computed for Figure \ref{MT-fig:fig4}.

>>>>

\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))],\label{eq:decomp:contrib_ATA_1}

@@

@@

\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))], \label{eq:decomp:contrib_ATA_2}

@@

Continuing, $\olep_i^{[E \nat C]}=\olep_i^{(EC)}-\olep_i^{[EC]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_2})
\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}

@@

\olep_j^{[EC]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\

@@

$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other. Continuing,
\olep_j^{[E \nat C]} &= \olep_j^{(EC)} - \olep_j^{[EC]} \\
&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:EChard}

@@

We now examine $\olep_i^{[EC]}$, which, by (\ref{eq:decomp:contrib_ATA_1}), equals 
\begin{equation}
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
\end{equation}
But $(\ddot{b}_i,\ddot{b}_j)$ is a bivariate normal random variable constructed to have the same marginals and 
covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
$(b_i^{(r)},b_j^{(r)})$, so $(\ddot{b}_i^{(l)},\ddot{b}_j^{(l)})$ is distributed in the same way as 
$(\ddot{b}_i^{(r)},\ddot{b}_j^{(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
of the choice between our left- and right-tail-associated noise cases. Now examining 
$\E[\ln(1-\delta+\delta\exp(b_i-b_j))]$, we define $\beta_1=-b_2+\mu_1+\mu_2$ and $\beta_2=-b_1+\mu_1+\mu_2$,
and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
$(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
$(b_1^{(l)},b_2^{(l)})$. But then we have 
\begin{align}
\E[\ln(1-\delta+\delta\exp(b_i^{(l)}-b_j^{(l)}))] &= \E[\ln(1-\delta+\delta\exp(-\beta_j^{(l)}+\mu_1+\mu_2-(-\beta_i^{(l)}+\mu_1+\mu_2)))] \\
&= \E[\ln(1-\delta+\delta\exp(\beta_i^{(l)}-\beta_j^{(l)}))] \\
&= \E[\ln(1-\delta+\delta\exp(b_i^{(r)}-b_j^{(r)}))], 
\end{align}
which completes the proof that $\olep_i^{[EC]}$ is independent of left- versus right-tail associated noise.
For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 
above, in this section, and note that all those terms depend only on marginals.

@@

Next, given values of the parameters $\delta$, $\mu_i - \mu_j$, and $\sigma$, proceed as follows to get estimates of the 
terms of the decomposition (\ref{eq:decomp_overall_IGR}).
\begin{enumerate}
\item Compute $\epsilon_i^0$ following (\ref{eq:decomp:baseline}), which is a simple formula of $\delta$ and 
$\mu_i - \mu_j$, not requiring use of the random samples described above. This gives an exact value of
$\epsilon_i^0$, not an estimate.
\item Estimate $\olep_i^E$, following (\ref{eq:decomp:contrib_of_var_E}), as 
$\widehat{\olep_i^E}=\mean_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0$, an estimate with 
standard error $\se \left( \widehat{\olep_i^E} \right) = \sd_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]/\sqrt{M}$,
where $\sd(\cdot)$ denotes sample standard deviation.
\item Estimate $\olep_i^C$, following (\ref{eq:decomp:contrib_of_var_C}), as 
$\widehat{\olep_i^C}=\mean_k [\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-\epsilon_i^0$, which has 
standard error $\se \left( \widehat{\olep_i^C} \right) = \sd_k[\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_i^\shp-\tilde{b}_j^\shp$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) is normally distributed
with mean $0$ and variance $2$, so we can estimate $\olep_i^{(E \shp C)}$, following (\ref{eq:decomp:contrib_of_EC_var_per_se}),
as $\widehat{\olep_i^{(E \shp C)}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0.$ The standard error of this expression is 
$\se \left( \widehat{\olep_i^{(E \shp C)}} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma \sqrt{2} u^{(k)}+\mu_i-\mu_j))-
\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$. 
\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_EC}), as
$\widehat{\olep_i^{[E \nat C]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$. The standard error of this expression is 
$\se \left( \widehat{\olep_i^{[E \nat C]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$
and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_ATA_2}), as 
$\widehat{\olep_i^{[EC]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]-
\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))].$ The standard
error of this expression is $\se \left( \widehat{\olep_i^{[EC]}} \right)=\sqrt{s_3^2+s_1^2}$, where
$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
\item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:base_IGR}), as 
$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))],$
which has standard error $s_3$ from above.
\end{enumerate}
It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
\widehat{\olep_i^{[E \nat C]}}+\widehat{\olep_i^{[EC]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
at the level of the estimators, as well

Next, proceed as follows to get estimates of the terms in the decomposition (\ref{eq:decomp_overall_rjbsi}). 
\begin{enumerate}
\item Use $\epsilon_j^0=0$, as in (\ref{eq:decomp:baseline_j}), and this is the exact value.
\item Estimate $\olep_j^E$, following (\ref{eq:decomp:contrib_of_var_E_j}), as
$\widehat{\olep_j^E}=\mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]-\epsilon_j^0$, an estimate
with standard error $\se \left( \widehat{\olep_j^E} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\olep_j^C$, following (\ref{eq:decomp:contrib_of_var_C_j}), as
$\widehat{\olep_j^C}=\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]-\epsilon_j^0$, which has 
standard error $\se \left( \widehat{\olep_j^C} \right) = \sd_k [\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_j$ and $\tilde{b}_j^\shp$ in (\ref{eq:decomp:j:EshpC}) are standard normally distributed and
independent, so $\tilde{b}_j^\shp - \tilde{b}_j$ is normally distributed with mean $0$ and variance $2$. So we can 
estimate $\olep_j^{(E \shp C)}$, following (\ref{eq:decomp:j:EshpC}), as 
$\widehat{\olep_j^{(E \shp C)}} = \mean_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\olep_j^E}-\widehat{\olep_j^C}-\epsilon_j^0$.
The standard error of this expression is $\se \left(  \widehat{\olep_j^{(E \shp C)}} \right) = 
\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))-\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\olep_j^{[E \nat C]}$, following (\ref{eq:decomp:j:EChard}), as $\widehat{\olep_j^{[E \nat C]}}=\mean_k [-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))],$ which has $\se \left( \widehat{\olep_j^{[E \nat C]}} \right) = \sd_k[-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))]/\sqrt{M}$.
\item Use $\olep_j^{[EC]}=0$, following (\ref{eq:decomp:j:EnatC}). This is exact.
\item Finally, use $\olr_{j \bs i} = 0$, again exact. 
\end{enumerate}
As for (\ref{eq:decomp_overall_IGR}), it is again easy to see that (\ref{eq:decomp_overall_rjbsi}) applies for the 
estimates as well as for the quantities being estimated.

@@

\item Estimate $\Delta_i^{[E \nat C]}$ as $\widehat{\Delta_i^{[E \nat C]}}=\widehat{\olep_i^{[E \nat C]}}-q_{ij}\widehat{\olep_j^{[E \nat C]}}=

@@

\item For $\Delta_i^{[EC]}$, we recall that $\olep_j^{[EC]}=0$, so $\Delta_i^{[EC]}=\olep_i^{[EC]}$. So we use
$\widehat{\olep_i^{[EC]}}$ for $\widehat{\Delta_i^{[EC]}}$.

@@

In Fig. \ref{MT-fig:fig4} we plot IGR , $\Delta_i^{[EC]}$, and IGR $-\Delta_i^{[EC]}$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.  
From section \ref{sect:math_simplifications}, 
\begin{align}
\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
&= 0,
\end{align}
and
\begin{align}
\olep_j^{[EC]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
&= 0. %\label{eq:decomp:j:EnatC}
\end{align}
Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[EC]}$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}$ are zero so 
\begin{equation}
\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\olr_{i \bs i}
\end{equation}
and 
\begin{equation}
\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}=\olep_i^{[EC]}
\end{equation}
Similarly for IGR $-\Delta_i^{[EC]}$, 
\begin{equation}
(\olr_{i \bs i}-q_{ij}\olr_{j \bs i})-(\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}) = \olr_{i \bs i}-\olep_i^{[EC]}.
\end{equation}
Thus $q_{ij}$ does not matter for anything computed for Fig. \ref{MT-fig:fig4}.
